---
title: "Sprint - Exploratory analysis"
author: "Carles Milà"
date: "2025-06-16"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.align = "center",
                      fig.width = 9)


library("dplyr")
library("sf")
library("GHRexplore")
library("here")
library("ggplot2")
library("lares")
library("cowplot")

# Env: Local
```

```{r Read data}
dengue_data <- read.csv(here("data", "processed", "weekly_data.csv")) |> 
  mutate(region_id = as.factor(region_id))
healthreg <- read_sf(here("data", "boundaries", "clean_hr.gpkg"))
states <- read_sf(here("data", "boundaries", "brazil_states.gpkg"))
```

```{r nino}
nino_ts <- dengue_data[c("date", "nino_year")]
nino_ts <- nino_ts[!duplicated(nino_ts),] 
nino_ts <- nino_ts |> 
  mutate(start = as.Date(date),
         end = start + 7,
         nino_year = factor(nino_year, levels = c("Nino", "Neutral", "Nina")))
```

# Spatial units

These are the 5 supra health regions we will use for visualizations, as well as maps for biome and koppen classifications.

```{r maps, fig.height=4}
p1 <- plot_map(dengue_data, var = "region_id", 
               time = "date", area = "regional_geocode",
               map = healthreg, map_area = "regional_geocode",
               by_year = FALSE)
p2 <- plot_map(dengue_data, var = "biome", 
               time = "date", area = "regional_geocode",
               map = healthreg, map_area = "regional_geocode",
               by_year = FALSE)
p3 <- plot_map(dengue_data, var = "koppen", 
               time = "date", area = "regional_geocode",
               map = healthreg, map_area = "regional_geocode", 
               by_year = FALSE)
plot_grid(p1, p2, p3, nrow = 1)
```

# Time series

## Incidence

```{r TS, fig.height=3}
dengue_data |> 
  plot_timeseries(type = "inc", time = "date", area = "regional_geocode",
                  var = "casos", pop = "pop", aggregate_space = "region_id",
                  title = "Incidence") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")
```

## Rolling mean 1

```{r TS-1, fig.height=3}
# tas
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "tas", pop = "pop", aggregate_space = "region_id",
                  title = "tas - 1 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# tas anomaly
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "tasan1", pop = "pop", aggregate_space = "region_id",
                  title = "tas anomaly - 1 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# prlr
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "prlr", pop = "pop", aggregate_space = "region_id",
                  title = "prlr - 1 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# spi
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "spi1", pop = "pop", aggregate_space = "region_id",
                  title = "SPI -1 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")
```

## Rolling mean 3

```{r TS-3, fig.height=3}
# tas
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "tas3", pop = "pop", aggregate_space = "region_id",
                  title = "tas - 3 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# tas anomaly
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "tasan3", pop = "pop", aggregate_space = "region_id",
                  title = "tas anomaly - 3 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# prlr
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "prlr3", pop = "pop", aggregate_space = "region_id",
                  title = "prlr - 3 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# spi
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "spi3", pop = "pop", aggregate_space = "region_id",
                  title = "SPI -3 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")
```

## Rolling mean 6

```{r TS-6, fig.height=3}
# tas
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "tas6", pop = "pop", aggregate_space = "region_id",
                  title = "tas - 6 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# tas anomaly
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "tasan6", pop = "pop", aggregate_space = "region_id",
                  title = "tas anomaly - 6 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# prlr
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "prlr6", pop = "pop", aggregate_space = "region_id",
                  title = "prlr - 6 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# spi
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "spi6", pop = "pop", aggregate_space = "region_id",
                  title = "SPI - 6 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")
```

## Rolling mean 12

```{r TS-12, fig.height=3}
# tas
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "tas12", pop = "pop", aggregate_space = "region_id",
                  title = "tas - 12 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# tas anomaly
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "tasan12", pop = "pop", aggregate_space = "region_id",
                  title = "tas anomaly - 12 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# prlr
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "prlr12", pop = "pop", aggregate_space = "region_id",
                  title = "prlr - 12 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")

# spi
dengue_data |> 
  plot_timeseries(type = "cov", time = "date", area = "regional_geocode",
                  var = "spi12", pop = "pop", aggregate_space = "region_id",
                  title = "SPI -12 (lag 0)") +
  geom_rect(data = nino_ts, 
            aes(xmin = start, xmax = end, ymin = I(0), ymax = I(350), 
                fill = nino_year), alpha = 0.1, inherit.aes = FALSE) +
  labs(fill = "Niño year")
```

# Incidence heat map

```{r heatmap, fig.height=6}
dengue_data |> 
  plot_heatmap(type = "inc", time = "date", area = "regional_geocode",
               var = "casos", pop = "pop", aggregate_space = "region_id")
```

# Incidence seasonality

```{r seasonality, fig.height=5}
dengue_data |> 
  plot_seasonality(type = "inc", time = "date", area = "regional_geocode",
                   var = "casos", pop = "pop", aggregate_space = "region_id")
```

# Correlation

The number before the dot indicates the number of months for the rolling average, the number after the dot indicates the lag. I show only lag0 for parsimony in the following graphs.

## tas, tasmin, tasmax, tasan

```{r cor tas, fig.height = 9}
datacor <- select(dengue_data, 
                  tas, tas3, tas6, tas12,
                  tasmin, tasmin3, tasmin6, tasmin12,
                  tasmax, tasmax3, tasmax6, tasmax12,
                  tasan1, tasan3, tasan6, tasan12)
plot_correlation(datacor, var = names(datacor), scale = 0.9)
```

## prlr, spi and spei

```{r cor spi}
datacor <- select(dengue_data, 
                  prlr, prlr3, prlr6, prlr12,
                  spi1, spi3, spi6, spi12,
                  spei1, spei3, spei6, spei12,)
plot_correlation(datacor, var = names(datacor), scale = 0.9)
```

## nino

```{r cor nino}
datacor <- select(dengue_data, 
                  nino, oni, nino6, nino12)
plot_correlation(datacor, var = names(datacor))
```

# Top raw correlations

```{r top, eval=FALSE}
datacor <- dengue_data
datacor$inc <- datacor$casos/datacor$pop
datacor <- datacor[,30:ncol(datacor)] |> 
  select(-ends_with("id"))
corr_var(datacor, var = "inc", top = 20)
```

# Spatial distribution

Case counts and incidence computed for the entire period.

```{r spatial distr, fig.width=8, fig.height=8}
p1 <- plot_map(dengue_data, var = "casos", type = "counts",
               area = "regional_geocode", time = "date",
               map = healthreg, map_area = "regional_geocode", 
               by_year = FALSE, transform = "log10p1") 
p2 <- plot_map(dengue_data, var = "casos", type = "inc",
               area = "regional_geocode", time = "date", pop = "pop",
               map = healthreg, map_area = "regional_geocode", 
               by_year = FALSE, transform = "log10p1", palette = "Green")
p3 <- plot_map(dengue_data, var = "koppen", 
               time = "date", area = "regional_geocode",
               map = healthreg, map_area = "regional_geocode", 
               by_year = FALSE)
plot_grid(p1, p2, p3, nrow = 2)
```
