---
title: "Univariate FE models"
author: "Carles Milà"
date: "2025-07-01"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.align = "center",
                      warning = FALSE,
                      message = FALSE)

library("GHRmodel")
library("dplyr")
library("forcats")
library("sf")
library("ggplot2")
library("cowplot")

# Env: Local
```

```{r baseline}
baseline <- read.csv("output/uni_fe/baseline_gof.csv")
baseline$model_id <- "intercept"
baseline$Type <- "Baseline"
```

# tas

## GOF

```{r tas gof, fig.width=9, fig.height=4}
# Read gof
tas_vars <- read.csv("output/uni_fe/tas_vars.csv")
tas_lin <- read.csv("output/uni_fe/tas_lin_gof.csv") |> 
  mutate(model_id = tas_vars$var, Type = "Linear") |> 
  select(-covariate_1)
tas_nl <- read.csv("output/uni_fe/tas_nl_gof.csv") |> 
  mutate(model_id = tas_vars$var, Type = "Non-linear") |> 
  select(-covariate_1)
tas_rep1 <- read.csv("output/uni_fe/tas_rep1_gof.csv") |> 
  mutate(model_id = tas_vars$var, Type = "Rep. region") |> 
  select(-covariate_1)
tas_rep2 <- read.csv("output/uni_fe/tas_rep2_gof.csv") |> 
  mutate(model_id = tas_vars$var, Type = "Rep. koppen") |> 
  select(-covariate_1)
tas_rep3 <- read.csv("output/uni_fe/tas_rep3_gof.csv") |> 
  mutate(model_id = tas_vars$var, Type = "Rep. biome") |> 
  select(-covariate_1)

# Delete outlier
tas_nl <- tas_nl[tas_nl$model_id != "tas6.l6",]

# Combine
tas_gof <- rbind(baseline, tas_lin, tas_nl, 
                 tas_rep1, tas_rep2, tas_rep3)

# DIC
ggplot(tas_gof, aes(x = model_id, y = dic, col = Type)) +
  geom_point() + 
  geom_hline(data = tas_gof[tas_gof$model_id=="intercept",],
             aes(yintercept = dic), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("DIC")

# WAIC
ggplot(tas_gof, aes(x = model_id, y = waic, col = Type)) +
  geom_point() + 
  geom_hline(data = tas_gof[tas_gof$model_id=="intercept",],
             aes(yintercept = waic), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("WAIC")


# LMS
ggplot(tas_gof, aes(x = model_id, y = lms, col = Type)) +
  geom_point() + 
  geom_hline(data = tas_gof[tas_gof$model_id=="intercept",],
             aes(yintercept = lms), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("LMS")

rm("tas_gof")
```

## Linear

```{r tas linear, fig.height=9}
mod_tas <- readRDS("output/uni_fe/tas_lin_mod.rds")
plot_coef_lin(mod_tas, mod_id = names(mod_tas$fixed)) +
  theme(legend.position = "none")
rm("mod_tas")
```

## Non-linear

Top 6 predictors only.

```{r tas nl}
mod_tas <- readRDS("output/uni_fe/tas_nl_mod.rds")
pvars <- c("tas12", "tas12.l1", "tasmax12", 
           "tasmax12.l1", "tasmax12.l2", "tasmax6")
p1 <-
  plot_coef_nl(mod_tas,
               mod_id = paste0("mod", which(tas_vars$var %in% pvars)[1:3]),
               pattern = "", xlab = "", ylab = "")
p2 <-
  plot_coef_nl(mod_tas,
               mod_id = paste0("mod", which(tas_vars$var %in% pvars)[4:6]),
               pattern = "", xlab = "", ylab = "")
plot_grid(p1, p2, ncol = 2)
rm("mod_tas")
```

## Rep - Region

Top 6 predictors only.

```{r tas rep1}
mod_tas <- readRDS("output/uni_fe/tas_rep1_mod.rds")
pvars <- c("tas12", "tas12.l1", "tasmax12", 
           "tasmax12.l1", "tasmax12.l2", "tasmax6")
pvars <- which(tas_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_tas, mod_id = paste0("mod", pvars[i]),
                                  name = "region_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_tas")
```

## Rep - Koppen

Top 6 predictors only.

```{r tas rep2}
mod_tas <- readRDS("output/uni_fe/tas_rep2_mod.rds")
pvars <- c("tas12", "tas12.l1", "tasmax12", 
           "tasmax12.l1", "tasmax12.l2", "tasmax6")
pvars <- which(tas_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_tas, mod_id = paste0("mod", pvars[i]),
                                  name = "koppen_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_tas")
```

## Rep - Biome

Top 6 predictors only.

```{r tas rep3}
mod_tas <- readRDS("output/uni_fe/tas_rep3_mod.rds")
pvars <- c("tas12", "tas12.l1", "tasmax12", 
           "tasmax12.l1", "tasmax12.l2", "tasmax6")
pvars <- which(tas_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_tas, mod_id = paste0("mod", pvars[i]),
                                  name = "biome_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_tas")
```

# tasan

## GOF

```{r tasan gof, fig.width=9, fig.height=4}
# Read gof
tasan_vars <- read.csv("output/uni_fe/tasan_vars.csv")
tasan_lin <- read.csv("output/uni_fe/tasan_lin_gof.csv") |> 
  mutate(model_id = tasan_vars$var, Type = "Linear") |> 
  select(-covariate_1)
tasan_nl <- read.csv("output/uni_fe/tasan_nl_gof.csv") |> 
  mutate(model_id = tasan_vars$var, Type = "Non-linear") |> 
  select(-covariate_1)
tasan_rep1 <- read.csv("output/uni_fe/tasan_rep1_gof.csv") |> 
  mutate(model_id = tasan_vars$var, Type = "Rep. region") |> 
  select(-covariate_1)
tasan_rep2 <- read.csv("output/uni_fe/tasan_rep2_gof.csv") |> 
  mutate(model_id = tasan_vars$var, Type = "Rep. koppen") |> 
  select(-covariate_1)
tasan_rep3 <- read.csv("output/uni_fe/tasan_rep3_gof.csv") |> 
  mutate(model_id = tasan_vars$var, Type = "Rep. biome") |> 
  select(-covariate_1)

# Delete outlier
tasan_nl <- tasan_nl[tasan_nl$model_id != "tasan6.l6",]

# Combine
tasan_gof <- rbind(baseline, tasan_lin, tasan_nl, 
                   tasan_rep1, tasan_rep2, tasan_rep3)

# DIC
ggplot(tasan_gof, aes(x = model_id, y = dic, col = Type)) +
  geom_point() + 
  geom_hline(data = tasan_gof[tasan_gof$model_id=="intercept",],
             aes(yintercept = dic), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("DIC")

# WAIC
ggplot(tasan_gof, aes(x = model_id, y = waic, col = Type)) +
  geom_point() + 
  geom_hline(data = tasan_gof[tasan_gof$model_id=="intercept",],
             aes(yintercept = waic), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("WAIC")


# LMS
ggplot(tasan_gof, aes(x = model_id, y = lms, col = Type)) +
  geom_point() + 
  geom_hline(data = tasan_gof[tasan_gof$model_id=="intercept",],
             aes(yintercept = lms), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("LMS")

rm("tasan_gof")
```

## Linear

```{r tasan linear}
mod_tasan <- readRDS("output/uni_fe/tasan_lin_mod.rds")
plot_coef_lin(mod_tasan, mod_id = names(mod_tasan$fixed)) +
  theme(legend.position = "none")
rm("mod_tasan")
```

## Non-linear

Top 6 predictors only.

```{r tasan nl}
mod_tasan <- readRDS("output/uni_fe/tasan_nl_mod.rds")
pvars <- c("tasan12", "tasan12.l1", "tasan12.l2", 
           "tasan6", "tasan6.l1", "tasan6.l2")
p1 <-
  plot_coef_nl(mod_tasan, 
               mod_id = paste0("mod", which(tasan_vars$var %in% pvars)[1:3]),
               pattern = "", xlab = "", ylab = "") 
p2 <-
  plot_coef_nl(mod_tasan, 
               mod_id = paste0("mod", which(tasan_vars$var %in% pvars)[4:6]),
               pattern = "", xlab = "", ylab = "") 
plot_grid(p1, p2, ncol = 2)
rm("mod_tasan")
```

## Rep - Region

Top 6 predictors only.

```{r tasan rep1}
mod_tasan <- readRDS("output/uni_fe/tasan_rep1_mod.rds")
pvars <- c("tasan12", "tasan12.l1", "tasan12.l2", 
           "tasan6", "tasan6.l1", "tasan6.l2")
pvars <- which(tasan_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_tasan, mod_id = paste0("mod", pvars[i]),
                                  name = "region_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_tasan")
```

## Rep - Koppen

Top 6 predictors only.

```{r tasan rep2}
mod_tasan <- readRDS("output/uni_fe/tasan_rep2_mod.rds")
pvars <- c("tasan12", "tasan12.l1", "tasan12.l2", 
           "tasan6", "tasan6.l1", "tasan6.l2")
pvars <- which(tasan_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_tasan, mod_id = paste0("mod", pvars[i]),
                                  name = "koppen_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_tasan")
```

## Rep - Biome

Top 6 predictors only.

```{r tasan rep3}
mod_tasan <- readRDS("output/uni_fe/tasan_rep3_mod.rds")
pvars <- c("tasan12", "tasan12.l1", "tasan12.l2", 
           "tasan6", "tasan6.l1", "tasan6.l2")
pvars <- which(tasan_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_tasan, mod_id = paste0("mod", pvars[i]),
                                  name = "biome_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_tasan")
```

# prlr

## GOF

```{r prlr gof, fig.width=9, fig.height=4}
# Read gof
prlr_vars <- read.csv("output/uni_fe/prlr_vars.csv")
prlr_lin <- read.csv("output/uni_fe/prlr_lin_gof.csv") |> 
  mutate(model_id = prlr_vars$var, Type = "Linear") |> 
  select(-covariate_1)
prlr_nl <- read.csv("output/uni_fe/prlr_nl_gof.csv") |> 
  mutate(model_id = prlr_vars$var, Type = "Non-linear") |> 
  select(-covariate_1)
prlr_rep1 <- read.csv("output/uni_fe/prlr_rep1_gof.csv") |> 
  mutate(model_id = prlr_vars$var, Type = "Rep. region") |> 
  select(-covariate_1)
prlr_rep2 <- read.csv("output/uni_fe/prlr_rep2_gof.csv") |> 
  mutate(model_id = prlr_vars$var, Type = "Rep. koppen") |> 
  select(-covariate_1)
prlr_rep3 <- read.csv("output/uni_fe/prlr_rep3_gof.csv") |> 
  mutate(model_id = prlr_vars$var, Type = "Rep. biome") |> 
  select(-covariate_1)

# Combine
prlr_gof <- rbind(baseline, prlr_lin, prlr_nl, 
                  prlr_rep1, prlr_rep2, prlr_rep3)

# DIC
ggplot(prlr_gof, aes(x = model_id, y = dic, col = Type)) +
  geom_point() + 
  geom_hline(data = prlr_gof[prlr_gof$model_id=="intercept",],
             aes(yintercept = dic), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("DIC")

# WAIC
ggplot(prlr_gof, aes(x = model_id, y = waic, col = Type)) +
  geom_point() + 
  geom_hline(data = prlr_gof[prlr_gof$model_id=="intercept",],
             aes(yintercept = waic), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("WAIC")


# LMS
ggplot(prlr_gof, aes(x = model_id, y = lms, col = Type)) +
  geom_point() + 
  geom_hline(data = prlr_gof[prlr_gof$model_id=="intercept",],
             aes(yintercept = lms), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("LMS")

rm("prlr_gof")
```

## Linear

```{r prlr linear}
mod_prlr <- readRDS("output/uni_fe/prlr_lin_mod.rds")
plot_coef_lin(mod_prlr, mod_id = names(mod_prlr$fixed)) +
  theme(legend.position = "none")
rm("mod_prlr")
```

## Non-linear

Top 6 predictors only.

```{r prlr nl}
mod_prlr <- readRDS("output/uni_fe/prlr_nl_mod.rds")
pvars <- c("prlr12.l1", "prlr12.l2", "prlr12.l3", 
           "prlr12.l4", "prlr12.l5", "prlr12.l6")
p1 <-
  plot_coef_nl(mod_prlr, 
               mod_id = paste0("mod", which(prlr_vars$var %in% pvars)[1:3]),
               pattern = "", xlab = "", ylab = "") 
p2 <-
  plot_coef_nl(mod_prlr, 
               mod_id = paste0("mod", which(prlr_vars$var %in% pvars)[4:6]),
               pattern = "", xlab = "", ylab = "") 
plot_grid(p1, p2, ncol = 2)
rm("mod_prlr")
```

## Rep - Region

Top 6 predictors only.

```{r prlr rep1}
mod_prlr <- readRDS("output/uni_fe/prlr_rep1_mod.rds")
pvars <- c("prlr12.l1", "prlr12.l2", "prlr12.l3", 
           "prlr12.l4", "prlr12.l5", "prlr12.l6")
pvars <- which(prlr_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_prlr, mod_id = paste0("mod", pvars[i]),
                                  name = "region_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_prlr")
```

## Rep - Koppen

Top 6 predictors only.

```{r prlr rep2}
mod_prlr <- readRDS("output/uni_fe/prlr_rep2_mod.rds")
pvars <- c("prlr12.l1", "prlr12.l2", "prlr12.l3", 
           "prlr12.l4", "prlr12.l5", "prlr12.l6")
pvars <- which(prlr_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_prlr, mod_id = paste0("mod", pvars[i]),
                                  name = "koppen_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_prlr")
```

## Rep - Biome

Top 6 predictors only.

```{r prlr rep3}
mod_prlr <- readRDS("output/uni_fe/prlr_rep3_mod.rds")
pvars <- c("prlr12.l1", "prlr12.l2", "prlr12.l3", 
           "prlr12.l4", "prlr12.l5", "prlr12.l6")
pvars <- which(prlr_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_prlr, mod_id = paste0("mod", pvars[i]),
                                  name = "biome_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_prlr")
```


# spi and spei

## GOF

```{r sp gof, fig.width=9, fig.height=4}
# Read gof
sp_vars <- read.csv("output/uni_fe/sp_vars.csv")
sp_lin <- read.csv("output/uni_fe/sp_lin_gof.csv") |> 
  mutate(model_id = sp_vars$var, Type = "Linear") |> 
  select(-covariate_1)
sp_nl <- read.csv("output/uni_fe/sp_nl_gof.csv") |> 
  mutate(model_id = sp_vars$var, Type = "Non-linear") |> 
  select(-covariate_1)
sp_rep1 <- read.csv("output/uni_fe/sp_rep1_gof.csv") |> 
  mutate(model_id = sp_vars$var, Type = "Rep. region") |> 
  select(-covariate_1)
sp_rep2 <- read.csv("output/uni_fe/sp_rep2_gof.csv") |> 
  mutate(model_id = sp_vars$var, Type = "Rep. koppen") |> 
  select(-covariate_1)
sp_rep3 <- read.csv("output/uni_fe/sp_rep3_gof.csv") |> 
  mutate(model_id = sp_vars$var, Type = "Rep. biome") |> 
  select(-covariate_1)

# Combine
sp_gof <- rbind(baseline, sp_lin, sp_nl, 
                sp_rep1, sp_rep2, sp_rep3)

# DIC
ggplot(sp_gof, aes(x = model_id, y = dic, col = Type)) +
  geom_point() + 
  geom_hline(data = sp_gof[sp_gof$model_id=="intercept",],
             aes(yintercept = dic), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("DIC")

# WAIC
ggplot(sp_gof, aes(x = model_id, y = waic, col = Type)) +
  geom_point() + 
  geom_hline(data = sp_gof[sp_gof$model_id=="intercept",],
             aes(yintercept = waic), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("WAIC")


# LMS
ggplot(sp_gof, aes(x = model_id, y = lms, col = Type)) +
  geom_point() + 
  geom_hline(data = sp_gof[sp_gof$model_id=="intercept",],
             aes(yintercept = lms), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("LMS")

rm("sp_gof")
```

## Linear

```{r sp linear, fig.height=5, fig.height=7}
mod_sp <- readRDS("output/uni_fe/sp_lin_mod.rds")
plot_coef_lin(mod_sp, mod_id = names(mod_sp$fixed)) +
  theme(legend.position = "none")
rm("mod_sp")
```

## Non-linear

Top 6 predictors only.

```{r sp nl}
mod_sp <- readRDS("output/uni_fe/sp_nl_mod.rds")
pvars <- c("spei12.l1", "spei12.l2", "spei12.l3", 
           "spi12.l1", "spi12.l2", "spi12.l3")
p1 <-
  plot_coef_nl(mod_sp, 
               mod_id = paste0("mod", which(sp_vars$var %in% pvars)[1:3]),
               pattern = "", xlab = "", ylab = "") 
p2 <-
  plot_coef_nl(mod_sp, 
               mod_id = paste0("mod", which(sp_vars$var %in% pvars)[4:6]),
               pattern = "", xlab = "", ylab = "") 
plot_grid(p1, p2, ncol = 2)
rm("mod_sp")
```

## Rep - Region

Top 6 predictors only.

```{r sp rep1}
mod_sp <- readRDS("output/uni_fe/sp_rep1_mod.rds")
pvars <- c("spei12.l1", "spei12.l2", "spei12.l3", 
           "spi12.l1", "spi12.l2", "spi12.l3")
pvars <- which(sp_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_sp, mod_id = paste0("mod", pvars[i]),
                                  name = "region_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_sp")
```

## Rep - Koppen

Top 6 predictors only.

```{r sp rep2}
mod_sp <- readRDS("output/uni_fe/sp_rep2_mod.rds")
pvars <- c("spei12.l1", "spei12.l2", "spei12.l3", 
           "spi12.l1", "spi12.l2", "spi12.l3")
pvars <- which(sp_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_sp, mod_id = paste0("mod", pvars[i]),
                                  name = "koppen_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_sp")
```

## Rep - Biome

Top 6 predictors only.

```{r sp rep3}
mod_sp <- readRDS("output/uni_fe/sp_rep3_mod.rds")
pvars <- c("spei12.l1", "spei12.l2", "spei12.l3", 
           "spi12.l1", "spi12.l2", "spi12.l3")
pvars <- which(sp_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_sp, mod_id = paste0("mod", pvars[i]),
                                  name = "biome_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_sp")
```



# nino

## GOF

```{r nino gof, fig.width=9, fig.height=4}
# Read gof
nino_vars <- read.csv("output/uni_fe/nino_vars.csv")
nino_lin <- read.csv("output/uni_fe/nino_lin_gof.csv") |> 
  mutate(model_id = nino_vars$var, Type = "Linear") |> 
  select(-covariate_1)
nino_nl <- read.csv("output/uni_fe/nino_nl_gof.csv") |> 
  mutate(model_id = nino_vars$var, Type = "Non-linear") |> 
  select(-covariate_1)
nino_rep1 <- read.csv("output/uni_fe/nino_rep1_gof.csv") |> 
  mutate(model_id = nino_vars$var, Type = "Rep. region") |> 
  select(-covariate_1)
nino_rep2 <- read.csv("output/uni_fe/nino_rep2_gof.csv") |> 
  mutate(model_id = nino_vars$var, Type = "Rep. koppen") |> 
  select(-covariate_1)
nino_rep3 <- read.csv("output/uni_fe/nino_rep3_gof.csv") |> 
  mutate(model_id = nino_vars$var, Type = "Rep. biome") |> 
  select(-covariate_1)

# Combine
nino_gof <- rbind(baseline, nino_lin, nino_nl, 
                  nino_rep1, nino_rep2, nino_rep3)

# DIC
ggplot(nino_gof, aes(x = model_id, y = dic, col = Type)) +
  geom_point() + 
  geom_hline(data = nino_gof[nino_gof$model_id=="intercept",],
             aes(yintercept = dic), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("DIC")

# WAIC
ggplot(nino_gof, aes(x = model_id, y = waic, col = Type)) +
  geom_point() + 
  geom_hline(data = nino_gof[nino_gof$model_id=="intercept",],
             aes(yintercept = waic), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("WAIC")


# LMS
ggplot(nino_gof, aes(x = model_id, y = lms, col = Type)) +
  geom_point() + 
  geom_hline(data = nino_gof[nino_gof$model_id=="intercept",],
             aes(yintercept = lms), alpha = 0.3, lty = 2) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom") +
  xlab("Predictor") + ylab("LMS")

rm("nino_gof")
```

## Linear

```{r nino linear}
mod_nino <- readRDS("output/uni_fe/nino_lin_mod.rds")
plot_coef_lin(mod_nino, mod_id = names(mod_nino$fixed)) +
  theme(legend.position = "none")
rm("mod_nino")
```

## Non-linear

Top 6 predictors only.

```{r nino nl}
mod_nino <- readRDS("output/uni_fe/nino_nl_mod.rds")
pvars <- c("nino12.l3", "nino6.l4", "nino6.l5", 
           "nino6.l6", "oni.l1", "oni.l6")
p1 <-
  plot_coef_nl(mod_nino, 
               mod_id = paste0("mod", which(nino_vars$var %in% pvars)[1:3]),
               pattern = "", xlab = "", ylab = "") 
p2 <-
  plot_coef_nl(mod_nino, 
               mod_id = paste0("mod", which(nino_vars$var %in% pvars)[4:6]),
               pattern = "", xlab = "", ylab = "") 
plot_grid(p1, p2, ncol = 2)
rm("mod_nino")
```

## Rep - Region

Top 6 predictors only.

```{r nino rep1}
mod_nino <- readRDS("output/uni_fe/nino_rep1_mod.rds")
pvars <- c("nino12", "nino6", "nino6.l1", 
           "nino6.l2", "oni.l2", "oni.l3")
pvars <- which(nino_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_nino, mod_id = paste0("mod", pvars[i]),
                                  name = "region_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_nino")
```

## Rep - Koppen

Top 6 predictors only.

```{r nino rep2}
mod_nino <- readRDS("output/uni_fe/nino_rep2_mod.rds")
pvars <- c("nino12", "nino6", "nino6.l1", 
           "nino6.l2", "oni.l2", "oni.l3")
pvars <- which(nino_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_nino, mod_id = paste0("mod", pvars[i]),
                                  name = "koppen_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_nino")
```

## Rep - Biome

Top 6 predictors only.

```{r nino rep3}
mod_nino <- readRDS("output/uni_fe/nino_rep3_mod.rds")
pvars <- c("nino12", "nino6", "nino6.l1", 
           "nino6.l2", "oni.l2", "oni.l3")
pvars <- which(nino_vars$var %in% pvars)
plist <- list()
for(i in 1:6){
  plist[[i]] <- plot_coef_varying(mod_nino, mod_id = paste0("mod", pvars[i]),
                                  name = "biome_id") 
}
plot_grid(plotlist = plist, ncol = 2)
rm("mod_nino")
```
