---
title: "ONI FE models"
author: "Carles Milà"
date: "2025-07-07"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.align = "center",
                      warning = FALSE,
                      message = FALSE)

library("GHRmodel")
library("dplyr")
library("sf")
library("ggplot2")
library("cowplot")

# Env: Local
# data <- read.csv("data/processed/weekly_data.csv") 
# koppen_labs <- levels(as.factor(data$koppen)); names(koppen_labs) <- 1:length(koppen_labs)
```


```{r modtab}
modtab <- read.csv("output/oni_fe/modtab.csv") |> 
  mutate(ID = paste0("mod", 1:n())) |> 
  select(ID, ONI, Model)
```

# GOF

Baseline model: LSL with weekly and yearly random effects

```{r gof, fig.width=7, fig.height=4}
modgof1 <- read.csv("output/oni_fe/spatial_norep_gof.csv") |> 
  mutate(SpatialRE = "Not replicated", 
         ONI = modtab$ONI, Model = modtab$Model)  |> 
  select(model_id, ONI, Model, SpatialRE, dic, waic, lms, mae)
modgof2 <- read.csv("output/oni_fe/spatial_rep_gof.csv") |> 
  mutate(SpatialRE = "Replicated",
         ONI = modtab$ONI, Model = modtab$Model)  |> 
  select(model_id, ONI, Model, SpatialRE, dic, waic, lms, mae)
modgof <- rbind(modgof1, modgof2)

# DIC
ggplot(modgof, aes(x = ONI, y = dic, col = Model, pch = SpatialRE)) +
  geom_hline(yintercept = modgof$dic[modgof$ONI == "Baseline"],
             col = "grey40", alpha = 0.4) +
  geom_point() +
  theme_bw()

# WAIC
ggplot(modgof, aes(x = ONI, y = waic, col = Model, pch = SpatialRE)) +
  geom_hline(yintercept = modgof$waic[modgof$ONI == "Baseline"],
             col = "grey40", alpha = 0.4) +
  geom_point() +
  theme_bw()

# LMS
ggplot(modgof, aes(x = ONI, y = lms, col = Model, pch = SpatialRE)) +
  geom_hline(yintercept = modgof$lms[modgof$ONI == "Baseline"],
             col = "grey40", alpha = 0.4) +
  geom_point() +
  theme_bw()

# MAE
ggplot(modgof, aes(x = ONI, y = mae, col = Model, pch = SpatialRE)) +
  geom_hline(yintercept = modgof$mae[modgof$ONI == "Baseline"],
             col = "grey40", alpha = 0.4) +
  geom_point() +
  theme_bw()
```

# Effects

```{r}
mod_norep <- readRDS("output/oni_fe/spatial_norep_mod.rds")
mod_rep <- readRDS("output/oni_fe/spatial_rep_mod.rds")
```

## Linear models

```{r effects linear, fig.width=6, fig.height=3}
p1 <- plot_coef_lin(mod_norep, mod_id = paste0("mod", 1:7)) + 
  theme(legend.position = "none") +
  ggtitle("Non-replicated spatial RE")
p2 <- plot_coef_lin(mod_rep, mod_id = paste0("mod", 1:7)) + 
  theme(legend.position = "none") +
  ggtitle("Replicated spatial RE")
plot_grid(p1, p2, nrow = 1)
```


## ENSO replicated by Koppen 

Non-replicated spatial RE

```{r}
p1 <- plot_coef_varying(mod_norep, mod_id = "mod20", name = "koppen_id8")
p2 <- plot_coef_varying(mod_norep, mod_id = "mod21", name = "koppen_id8")
p3 <- plot_coef_varying(mod_norep, mod_id = "mod22", name = "koppen_id8")
p4 <- plot_coef_varying(mod_norep, mod_id = "mod23", name = "koppen_id8")
p5 <- plot_coef_varying(mod_norep, mod_id = "mod24", name = "koppen_id8")
p6 <- plot_coef_varying(mod_norep, mod_id = "mod25", name = "koppen_id8")
plot_grid(p1, p2, p3, p4, p5, p6, nrow = 2)
```

Replicated spatial RE

```{r}
p1 <- plot_coef_varying(mod_rep, mod_id = "mod20", name = "koppen_id8")
p2 <- plot_coef_varying(mod_rep, mod_id = "mod21", name = "koppen_id8")
p3 <- plot_coef_varying(mod_rep, mod_id = "mod22", name = "koppen_id8")
p4 <- plot_coef_varying(mod_rep, mod_id = "mod23", name = "koppen_id8")
p5 <- plot_coef_varying(mod_rep, mod_id = "mod24", name = "koppen_id8")
p6 <- plot_coef_varying(mod_rep, mod_id = "mod25", name = "koppen_id8")
plot_grid(p1, p2, p3, p4, p5, p6, nrow = 2)
```

## ENSO replicated by region

Non-replicated spatial RE

```{r}
p1 <- plot_coef_varying(mod_norep, mod_id = "mod26", name = "region_id")
p2 <- plot_coef_varying(mod_norep, mod_id = "mod27", name = "region_id")
p3 <- plot_coef_varying(mod_norep, mod_id = "mod28", name = "region_id")
p4 <- plot_coef_varying(mod_norep, mod_id = "mod29", name = "region_id")
p5 <- plot_coef_varying(mod_norep, mod_id = "mod30", name = "region_id")
p6 <- plot_coef_varying(mod_norep, mod_id = "mod31", name = "region_id")
plot_grid(p1, p2, p3, p4, p5, p6, nrow = 2)
```

Replicated spatial RE

```{r}
p1 <- plot_coef_varying(mod_rep, mod_id = "mod26", name = "region_id")
p2 <- plot_coef_varying(mod_rep, mod_id = "mod27", name = "region_id")
p3 <- plot_coef_varying(mod_rep, mod_id = "mod28", name = "region_id")
p4 <- plot_coef_varying(mod_rep, mod_id = "mod29", name = "region_id")
p5 <- plot_coef_varying(mod_rep, mod_id = "mod30", name = "region_id")
p6 <- plot_coef_varying(mod_rep, mod_id = "mod31", name = "region_id")
plot_grid(p1, p2, p3, p4, p5, p6, nrow = 2)
```

## ENSO replicated by biome

Non-replicated spatial RE

```{r}
p1 <- plot_coef_varying(mod_norep, mod_id = "mod32", name = "biome_id")
p2 <- plot_coef_varying(mod_norep, mod_id = "mod33", name = "biome_id")
p3 <- plot_coef_varying(mod_norep, mod_id = "mod34", name = "biome_id")
p4 <- plot_coef_varying(mod_norep, mod_id = "mod35", name = "biome_id")
p5 <- plot_coef_varying(mod_norep, mod_id = "mod36", name = "biome_id")
p6 <- plot_coef_varying(mod_norep, mod_id = "mod37", name = "biome_id")
plot_grid(p1, p2, p3, p4, p5, p6, nrow = 2)
```

Replicated spatial RE

```{r}
p1 <- plot_coef_varying(mod_rep, mod_id = "mod32", name = "biome_id")
p2 <- plot_coef_varying(mod_rep, mod_id = "mod33", name = "biome_id")
p3 <- plot_coef_varying(mod_rep, mod_id = "mod34", name = "biome_id")
p4 <- plot_coef_varying(mod_rep, mod_id = "mod35", name = "biome_id")
p5 <- plot_coef_varying(mod_rep, mod_id = "mod36", name = "biome_id")
p6 <- plot_coef_varying(mod_rep, mod_id = "mod37", name = "biome_id")
plot_grid(p1, p2, p3, p4, p5, p6, nrow = 2)
```

## ENSO non-linear (equal)

Non-replicated spatial RE

```{r, fig.width=8, fig.height=4}
p1 <- plot_coef_nl(mod_norep, mod_id = "mod8")
p2 <- plot_coef_nl(mod_norep, mod_id = "mod9")
p3 <- plot_coef_nl(mod_norep, mod_id = "mod10")
p4 <- plot_coef_nl(mod_norep, mod_id = "mod11")
p5 <- plot_coef_nl(mod_norep, mod_id = "mod12")
p6 <- plot_coef_nl(mod_norep, mod_id = "mod13")
cowplot::plot_grid(p1, p2, p3, p4, p5, p6, nrow = 2)
```

Replicated spatial RE

```{r, fig.width=8, fig.height=4}
p1 <- plot_coef_nl(mod_rep, mod_id = "mod8")
p2 <- plot_coef_nl(mod_rep, mod_id = "mod9")
p3 <- plot_coef_nl(mod_rep, mod_id = "mod10")
p4 <- plot_coef_nl(mod_rep, mod_id = "mod11")
p5 <- plot_coef_nl(mod_rep, mod_id = "mod12")
p6 <- plot_coef_nl(mod_rep, mod_id = "mod13")
cowplot::plot_grid(p1, p2, p3, p4, p5, p6, nrow = 2)
```

## ENSO non-linear (quantile)

Non-replicated spatial RE

```{r, fig.width=8, fig.height=4}
p1 <- plot_coef_nl(mod_norep, mod_id = "mod14")
p2 <- plot_coef_nl(mod_norep, mod_id = "mod15")
p3 <- plot_coef_nl(mod_norep, mod_id = "mod16")
p4 <- plot_coef_nl(mod_norep, mod_id = "mod17")
p5 <- plot_coef_nl(mod_norep, mod_id = "mod18")
p6 <- plot_coef_nl(mod_norep, mod_id = "mod19")
cowplot::plot_grid(p1, p2, p3, p4, p5, p6, nrow = 2)
```

Replicated spatial RE

```{r, fig.width=8, fig.height=4}
p1 <- plot_coef_nl(mod_rep, mod_id = "mod14")
p2 <- plot_coef_nl(mod_rep, mod_id = "mod15")
p3 <- plot_coef_nl(mod_rep, mod_id = "mod16")
p4 <- plot_coef_nl(mod_rep, mod_id = "mod17")
p5 <- plot_coef_nl(mod_rep, mod_id = "mod18")
p6 <- plot_coef_nl(mod_rep, mod_id = "mod19")
cowplot::plot_grid(p1, p2, p3, p4, p5, p6, nrow = 2)
```